Formalizing Stack Safety as a Security Property

### 2021-09-30 submission4/main.tex
Lots of gains with space hacking
wibble
12 pages (and some comments)

~$\cdots$
\vspace{.5em}
\\
3 &
\memoryaddrs{8em}
\memory{3}{\unsealc}
~$\cdots$
    \MemoryLabel{-15em}{0.75em}{42}
    \vspace{.5em}
\\
12 &
\memoryaddrs{16em}
\memory{1}{\mainsealc}[\sealdesc{0}]%
\memory{1}{\retptrc}[\retptrdesc]%
\memory{1}{\unsealc}[\unsealdesc]
~$\cdots$
\MemoryLabel{-16em}{0.75em}{42}
\MemoryLabel{-7.5em}{0.75em}{42}
\vspace{.5em}
\\
13 &
\memoryaddrs{16em}
\memory{1}{\mainsealc}
\memory{1}{\retptrc}
\memory{1}{\unsealc}
~$\cdots$
\MemoryLabel{-15em}{0.75em}{\bf 0}
\vspace{.5em}
\end{tabular}
\end{center}
\vspace{-4ex}
\end{subfigure}
\caption{Integrity and confidentiality violations
\ifaftersubmission\bcp{I wonder if we
    could make the figures easier to read by putting very light, differently
  colored backgrounds under the different parts} \sna{I like the minimalism of this, but the code is now set up to support playing with it}\fi}
\label{fig:int1}
\end{figure*}

@@

\caption{Noninterference, violated}
\label{fig:conf1}
\end{figure}

@@

\begin{figure}
%\begin{center}
\begin{subfigure}[t]{.28\columnwidth}
  {\small
\begin{lstlisting}
int main() {
  int x;
  x = f();
  return g();
}
\end{lstlisting}
  }
\end{subfigure}
\begin{subfigure}[t]{.27\columnwidth}
  {\small
\begin{lstlisting}[backgroundcolor=\color{lgray}]
int f() {
  int y = 5;
  return y;
}
\end{lstlisting}
  }
\end{subfigure}
\begin{subfigure}[t]{.32\columnwidth}
  {\small
\begin{lstlisting}
int g() {
  int z;
  return z;
}
\end{lstlisting}
  }
\end{subfigure}
\caption{Another confidentiality violation}
\label{fig:conf2}
\end{figure}

@@

Finally, we formalize WBCF as a simple property of the behavior of returns.
\definition
A system enjoys \textit{\textbf{well-bracketed control flow}} when, for every reachable state
\(\mpcstatename\) where \(\pi_\context(\mpcstatename) = (\underscore,\rts)\), and
\(\mpcstatename \stepsto \mpcstatename'\) where \(\pi_\context(\mpcstatename') = (\underscore,\rts')\),
if \(|\rts'| < |\rts|\), then \(|\rts'| = |\rts|-1\).

@@

\paragraph*{Conservative Stack-Safety Enforcement Mechanism}
%
The enforcement policy of Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} relies
on a programmable, tag-based reference monitor that runs alongside the
program,
allowing fine-grained manipulation of metadata tags to encode so-called
\emph{micro-policies}~\cite{pump_oakland2015}.
Every value in
memory and registers (including the $\PCname$)
is enriched with an abstract metadata tag, which can represent
arbitrary information about the value. A micro-policy is then defined as a
set of tags and a
collection of software-defined rules.
At each step of the machine, the relevant rule is applied to the tags on
the instruction's inputs ($\PCname$, registers, memory) and on the
instruction opcode itself, producing
one of two outcomes: either the instruction is permitted to execute (and
generates tags for the result of the operation and the new $\PCname$), or
the machine fail-stops with a policy violation.
Existing work \cite{TestingNI:ICFP,pump_oakland2015, DBLP:conf/sp/RoesslerD18}
%\bcp{additional citations: the noninterference paper, Nick and Andre's papers, ...}
has shown that a wide range of
micro-policies can be defined using this scheme.
Efficient execution of these micro-policies relies on hardware acceleration,
such as the PUMP architecture~\cite{pump:asplos2015}.

@@

These two properties give us
{\em Observable Stack Safety}. This is a weaker property
than the stepwise one, and also much more extensional.
We can think of the stepwise properties as instantiating
the observation model with a view of the entire state, and
thus as one end of a spectrum that can become arbitrarily permissive.
Therefore, if we had to choose a single property to represent stack safety,
it would be the observational formalization.

@@

\medskip
\paragraph*{Parameters Passed on the Stack}
%
Up to this point we have ignored function arguments, by assuming that they are placed
in registers (which are implicitly shared by all functions).
We now show how to implement two kinds of protected
sharing of stack contents: passing by value on the stack, and passing by reference. We first explore
some examples and then extend our formalization.

@@

i) Our definition captures the intuition that {\em
  every} caller should be protected from its callees.  By contrast,
  Skorstengaard et al. \cite{SkorstengaardSTKJFP} describe a class of ``unreasonable'' elements
  that are not protected, because they are capable of dynamic behavior that can compromise
  their own security. Consider a function that, when called with the wrong
  inputs, behaves unreasonably. Skorstengaard et al. \cite{SkorstengaardSTKJFP} would not consider this a
  {\em reasonable element} (as defined in their section 4.3.3), as a reasonable
  element is always reasonable. In our setting, individual
  well-behaved activations of the buggy function would be protected.

@@

Skorstengaard et al. \cite{SkorstengaardSTKJFP} propose a useful list of five criteria for judging
candidate stack safety properties.
%% : they should be (1) intuitive; (2)
%% useful for reasoning; (3) reusable in secure compiler chains; (4)
%% arguably ``complete'' with respect to informal notions of stack
%% safety; and (5) potentially scalable to richer settings such as
%% dynamic code generation or multithreading.
%
(1) {\em Intuitiveness}. Our property definitions are at least as
intuitive as theirs, while being phrased in interestingly different terms.
We use the language of formal security
properties to apply well-established concepts and formal techniques
of noninterference, and we define our properties in terms of
traces of memory and register states, rather than in terms of an
ad-hoc abstract machine.
Which approach is ``better'' is probably a matter of taste, but ours
has at least the merit of explicitly relating stack safety to existing
notions of security.
%
(2) {\em Usefulness} of the safety criterion
for carrying out formal proofs (about individual programs). We agree that this is an important consideration;
demonstrating it for our proposal remains future work, in which we
plan to apply our properties to justifying a low-level program logic.
%
(3) {\em Reusability} in secure compilation chains.
StkTokens is specifically set up to address this,
while it is not a goal of our work.
%
(4) {\em Comprehensiveness} in capturing programmers' informal
understanding of stack safety. We believe that our support
for dynamically changing notion of adversaries, as outlined in point (i) above,
captures a better definition of safety than LSE; moreover, our observational
properties capture an even richer and more extensional notion of stack safety
that can be applied to a broader range of enforcement mechanisms and code features.
%
(5) {\em Broad applicability} to richer settings. We have demonstrated that our approach extends to
coroutines, a first step towards multithreading. Dynamic code generation is
more challenging, but StkTokens does not yet handle this either.

Finally, we would propose one more criterion: reusability in evaluating multiple machines and
enforcement mechanisms. They have used their approach to give strong validation
to one machine/enforcement mechanism, a linear-capability machine,
via formal proof. We have used ours to obtain a weaker, testing-based validation of
two different mechanisms on a single machine, and will target a capability machine
next as future work.

@@

We add a rule to handle {\sc SetJump} annotations, and modify the {\sc Call} rule to insert {\sc CR} labels into the
return stack.
\judgment[]
         {\(\codemap ~ (\mach ~ \PCname) = {\sc SetJump}\)}
         {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\text{\sc EX}, \retcon ~ \mach)::\rts\)}
\judgmenttwobrlong[]
              {\(\codemap ~ (\mach ~ \PCname) = \callmap\)}
              {\(\components = \{\component | \sealcon ~ \mach[\component]\}\)}
              {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}]\)}
              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\text{\sc CR}, \retcon ~ \mach)::\rts\)}
The return rules operate identically regardless of the target labels.

>>>>

~$\cdots$
\vspace{.5em}
\\
3 &
\memoryaddrs{8em}
\memory{3}{\unsealc}
~$\cdots$
    \MemoryLabel{-15em}{0.75em}{42}
    \vspace{.5em}
\\
12 &
\memoryaddrs{16em}
\memory{1}{\mainsealc}[\sealdesc{0}]%
\memory{1}{\retptrc}[\retptrdesc]%
\memory{1}{\unsealc}[\unsealdesc]
~$\cdots$
\MemoryLabel{-16em}{0.75em}{42}
\MemoryLabel{-7.5em}{0.75em}{42}
\vspace{.5em}
\\
13 &
\memoryaddrs{16em}
\memory{1}{\mainsealc}
\memory{1}{\retptrc}
\memory{1}{\unsealc}
~$\cdots$
\MemoryLabel{-15em}{0.75em}{\bf 0}
\vspace{.5em}
\end{tabular}
\end{center}
\vspace{-4ex}
\end{subfigure}
\caption{Integrity and confidentiality violations
\ifaftersubmission\bcp{I wonder if we
    could make the figures easier to read by putting very light, differently
    colored backgrounds under the different parts} \sna{I like the minimalism of this, but the code is now set up to support playing with it}\fi}
\vspace*{-1em}
\label{fig:int1}
\end{figure*}

@@

\caption{Noninterference, violated}
\vspace*{-1em}
\label{fig:conf1}
\end{figure}

@@

\begin{figure}
%\begin{center}
\begin{subfigure}[t]{.28\columnwidth}
  {\small
\begin{lstlisting}
int main() {
  int x;
  x = f();
  return g();
}
\end{lstlisting}
  }
\end{subfigure}
\begin{subfigure}[t]{.27\columnwidth}
  {\small
\begin{lstlisting}[backgroundcolor=\color{lgray}]
int f() {
  int y = 5;
  return y;
}
\end{lstlisting}
  }
\end{subfigure}
\begin{subfigure}[t]{.32\columnwidth}
  {\small
\begin{lstlisting}
int g() {
  int z;
  return z;
}
\end{lstlisting}
  }
\end{subfigure}
\caption{Another confidentiality violation}
\vspace*{-1em}
\label{fig:conf2}
\end{figure}

@@

Finally, WBCF is a \ifspace simple \fi property of the behavior of returns.
\definition
A system enjoys \textit{\textbf{well-bracketed control flow}} when, for every reachable state
\(\mpcstatename\) where \(\pi_\context(\mpcstatename) = (\underscore,\rts)\), and
\(\mpcstatename \stepsto \mpcstatename'\) where \(\pi_\context(\mpcstatename') = (\underscore,\rts')\),
if \(|\rts'| < |\rts|\), then \(|\rts'| = |\rts|-1\).

@@

\paragraph*{Conservative Stack-Safety Enforcement Mechanism}
%
The enforcement policy of Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} relies
on a programmable, tag-based reference monitor that runs alongside the
program,
allowing fine-grained manipulation of metadata tags to encode so-called
\emph{micro-policies}~\cite{pump_oakland2015}.
Every value in
memory and registers (including the $\PCname$)
is enriched with an abstract metadata tag, which can represent
arbitrary information about the value. A micro-policy is then defined as a
set of tags and a
collection of software-defined rules.
On each machine step, the relevant rule is applied to the tags on
the instruction's inputs ($\PCname$, registers, memory) and on the
\ifspace instruction \fi opcode \ifspace itself\fi, producing
one of two outcomes: either the instruction is permitted to execute (and
generates tags for the result of the operation and the new $\PCname$), or
the machine fail-stops with a policy violation.
Existing work \cite{TestingNI:ICFP,pump_oakland2015, DBLP:conf/sp/RoesslerD18}
%\bcp{additional citations: the noninterference paper, Nick and Andre's papers, ...}
has shown that a wide range of
micro-policies can be defined using this scheme.
Efficient execution of these micro-policies relies on hardware acceleration,
such as the PUMP architecture~\cite{pump:asplos2015}.

@@

These two properties give us
{\em Observable Stack Safety}. This is a weaker property
than the stepwise one, and also much more extensional.
We can think of the stepwise properties as instantiating
the observation model with a view of the entire state---one end of a
spectrum that can become arbitrarily permissive.
Thus, if we had to choose a single property to represent stack safety,
it would be the observational one.

@@

\medskip
\paragraph*{Parameters Passed on the Stack}
%
Up to this point we have ignored function arguments, by assuming that they are placed
in registers (which are implicitly shared by all functions).
We now show how to implement two kinds of protected
sharing of stack contents: passing by value on the stack, and passing by
reference.
\ifspace
We first explore
some examples and then extend our formalization.
\fi

@@

i) Our definition captures the intuition that {\em
  every} caller should be protected from its callees.  By contrast,
  Skorstengaard et al. \cite{SkorstengaardSTKJFP} describe a class of ``unreasonable'' elements
  that are not protected, because they are capable of compromising
  their own security. Consider a function that, when called with the wrong
  inputs, behaves unreasonably. Skorstengaard et al. \cite{SkorstengaardSTKJFP} would not consider this a
  {\em reasonable element} (as defined in their section 4.3.3), as a reasonable
  element is always reasonable. In our setting, individual
  well-behaved activations of the buggy function would be protected.

@@

Skorstengaard et al. \cite{SkorstengaardSTKJFP} propose a useful list of five criteria for judging
candidate stack safety properties.
%% : they should be (1) intuitive; (2)
%% useful for reasoning; (3) reusable in secure compiler chains; (4)
%% arguably ``complete'' with respect to informal notions of stack
%% safety; and (5) potentially scalable to richer settings such as
%% dynamic code generation or multithreading.
%
In terms of (1) {\em intuitiveness}, we contend that our property
definitions are at least as
intuitive as theirs, though phrased in rather different terms.
We use the language of formal security
properties to apply well-established concepts \ifspace and formal techniques \fi
of noninterference, and we phrase our properties in terms of
traces of memory and register states, rather than an
ad-hoc abstract machine.
Which approach is ``better'' is probably a matter of taste, but ours
has \ifspace at least \fi the merit of explicitly relating stack safety to existing
notions of security.
%
Demonstrating (2) {\em usefulness} of the safety criterion
for carrying out formal proofs (about individual programs) for our proposal
remains future work; concretely, we
hope to apply our properties to justifying a low-level program logic.
%
(3) StkTokens is specifically set up to address {\em reusability} in secure
compilation chains, while this is not a goal of our work.\bcp{Can we explain
what it means??}
%
(4) Regarding {\em comprehensiveness} in capturing programmers' informal
understanding of stack safety, we believe that our support
for a dynamically changing notion of ``adversary,'' as outlined in point (i) above,
captures an even better definition of safety than LSE; moreover, our observational
properties are richer and\bcp{cut "richar and"??} more extensional, and thus
applicable to a broader range of enforcement mechanisms and code features.
%
(5) Finally, regarding {\em broad applicability}, we have demonstrated that
\bcp{unlike StkTokens}
our approach extends to
coroutines, a first step towards multithreading. Dynamic code generation is
more challenging, but StkTokens does not yet handle this either.

We would add one more criterion: reusability across machine models and
enforcement mechanisms. StkTokens has been used to validatate
one machine/enforcement mechanism---a linear-capability machine---via formal
proof. We have used ours to obtain a weaker (testing-based) validation of
two different mechanisms on a single machine; we hope to target a capability machine
next as future work.

@@

We add a rule to handle {\sc SetJump} annotations, and modify the {\sc Call} rule to insert {\sc CR} labels into the
return stack.

\judgment[]
         {\(\codemap ~ (\mach ~ \PCname) = {\sc SetJump}\)}
         {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\text{\sc EX}, \retcon ~ \mach)::\rts\)}
\judgmenttwobrlong[]
              {\(\codemap ~ (\mach ~ \PCname) = \callmap\)}
              {\(\components = \{\component | \sealcon ~ \mach[\component]\}\)}
              {\(\domMap' = \domMap[\components \mapsto \sealed{|\rts|}]\)}
              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\text{\sc CR}, \retcon ~ \mach)::\rts\)}
The return rules operate identically regardless of the target labels.
